Viscous Erosion in Porous Media

### 2018-03-16 docs/methods/erosion_methods.tex
NM

Figure \ref{01bodseq} shows the erosion of a single body, numerically simulated with $N_\iin = 1024$ discretization points and shown at three evenly spaced times. Here, color represents the vorticity of the surrounding flow. When evaluated on the boundary, vorticity corresponds to the shear stress and thus the local erosion rate. As seen in the left-most snapshot, the body is initially circular and has the strongest vorticity above and below it. The flow is for-aft symmetric as a result of the time-reversibility of the Stokes equations, with vorticity vanishing at the body's front and rear stagnation points. 
At the next instance, the body has lost material due to erosion, but remains for-aft symmetric due to the flow symmetry. The greatest material loss has occurred at the body's top and bottom, where shear is strongest. Interestingly, the body has developed well-defined corners at its front and rear stagnation points. This geometric feature is a consequence of the erosion law, which depends on the {\em absolute} shear stress: wherever the stress vanishes, $\atau$ exhibits a corner, which manifests as a corner in the geometry. As shown in the third snapshot, this corner persists as the bodies becomes even more slender and continues to shrink. In this simulation, the body completely vanishes at a final time of $t_f = 1.79 \times 10^{-2}$.
\todo[inline]{Is this the notation we want to use for scientific
notation? Could instead use $1.79 \times 10^{-2}$? \\
Fine by me (Nick). What about for tables (e.g. table 4)? For tables, I am more used to seeing the 1.79E-2 notation. How about you?}

To better understand the shape evolution, Fig.~\ref{shrink_intface}(a) shows several superimposed interfaces at evenly spaced times. The interfaces are color-coded by normalized time $t/t_f$. The body seems to converge to a shape that resembles an American football --- a fairly slender, for-aft symmetric shape with corners at the front and back.
Fig.~\ref{shrink_intface}(b) shows how the absolute shear stress, $\atau$, is distributed along each of these interfaces. The normalized arclength $s/L$ runs counterclockwise around the body, with $s/L = 0$ corresponding to the rear stagnation point and $s/L = 0.5$ the front stagnation point. At early times (yellow to orange), the stress is distributed non-uniformly, with highest values at the top and bottom of the body. As time proceeds, though, the stress becomes more evenly distributed along the body's surface (red). As the body nears the final stages of erosion (dark red), the stress actually becomes slightly higher near the stagnation points. A close look reveals that the stress always vanishes {\em at} the stagnation points, but in a region that becomes increasingly narrow.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{./figs/shrink_intface.pdf}
\caption{
Shape and stress evolution: (a) Interfaces of the eroding body from Fig.~\ref{01bodseq} at evenly-spaced time intervals. Color indicates normalized time, $t/t_f$, where $t_f = 1.79 \times 10^{-2}$ is the time at which the body completely vanishes. (b) The distribution of absolute shear stress, $\atau$, along each of these interfaces. Normalized arclength, $s/L$, runs counterclockwise from the rear stagnation point. The stress initially varies over the body's surface, but becomes more uniform over time.}
\label{shrink_intface}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure from datafile 01circ1024aa;
% Uses 01circ1024.in with epsfac = 10, sigfac = 10, dt = 1e-6, fixarea = 0, fixpdrop = 0

>>>>

Figure \ref{01bodseq} shows the erosion of a single body, numerically simulated with $N_\iin = 1024$ discretization points and shown at three evenly spaced times. Here, color represents the vorticity of the surrounding flow. When evaluated on the boundary, vorticity corresponds to the shear stress and thus the local erosion rate. As seen in the left-most snapshot, the body is initially circular and has the strongest vorticity above and below it. The flow is for-aft symmetric as a result of the time-reversibility of the Stokes equations, with vorticity vanishing at the body's front and rear stagnation points. 
At the next instance, the body has lost material due to erosion, but remains for-aft symmetric due to the flow symmetry. The greatest material loss has occurred at the body's top and bottom, where shear is strongest. Interestingly, the body has developed well-defined corners at its front and rear stagnation points. This geometric feature is a consequence of the erosion law, which depends on the {\em absolute} shear stress: wherever the stress vanishes, $\atau$ exhibits a corner, which manifests as a corner in the geometry. As shown in the third snapshot, this corner persists as the bodies becomes even more slender and continues to shrink.
\todo[inline]{Is this the notation we want to use for scientific
notation? Could instead use $1.79 \times 10^{-2}$? \\
Fine by me (Nick). What about for tables (e.g. table 4)? For tables, I am more used to seeing the 1.79E-2 notation. How about you?}

To better understand the shape evolution, Fig.~\ref{shrink_intface}(a) shows several superimposed interfaces at evenly spaced times. The interfaces are color-coded by normalized time $t/t_f$, where $t_f = 1.79 \times 10^{-2}$ is the time at which the body vanishes. The body converges to a fairly slender, for-aft symmetric shape with corners at its front and back --- somewhat reminiscent of an American football.
Fig.~\ref{shrink_intface}(b) shows how the absolute shear stress, $\atau$, is distributed along each of these interfaces. The normalized arclength $s/L$ runs counterclockwise around the body, with $s/L = $ 0 and 0.5 corresponding to the rear and front stagnation points respectively. At early times (yellow to orange), the stress is highly variables, with highest values at the top and bottom of the body. As time proceeds, though, the stress becomes more evenly distributed along the body's surface (red). As the body nears the final stages of erosion (dark red), the stress actually becomes slightly higher near the stagnation points. A close look reveals that the stress always vanishes {\em at} the stagnation points, but in a region that becomes increasingly narrow.

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.8 \textwidth]{./figs/shrink_intface.pdf}
\caption{
Shape and stress evolution: (a) Interfaces of the eroding body from Fig.~\ref{01bodseq} shown at evenly-spaced time intervals. Color indicates normalized time, $t/t_f$, where $t_f = 1.79 \times 10^{-2}$ is the time at which the body completely vanishes. (b) The distribution of absolute shear stress, $\atau$, along each of these interfaces. Normalized arclength, $s/L$, runs counterclockwise from the rear stagnation point. The stress initially varies over the body's surface, but becomes more uniform over time.}
\label{shrink_intface}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure from datafile 01circ1024aa;
% Uses 01circ1024.in with epsfac = 10, sigfac = 10, dt = 1e-6, fixarea = 0, fixpdrop = 0
